Variants on the Flight and Maintenance Planning problem

### 2019-04-18 latex/notes/model/model.tex
new version of bounds and model documents.

        \begin{tabular}{p{15mm}p{140mm}}
            $i \in \mathcal{I}$     &  aircraft. \\
            $t \in \mathcal{T}$     &  time periods included in the planning horizon. \\
            $j \in \mathcal{J}$     &  missions. \\
            $k \in \mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
        \end{tabular}

@@

        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
            $N_t$               & number of aircraft known to be in maintenance in period $t$. & aircraft. \\
            $N^{Clust}_{kt}$    & number of aircraft in cluster $k$ known to be in maintenance in period $t$. & aircraft. \\
            $A^{Clust}_{kt}$    & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$. & aircraft. \\
            $H^{Clust}_{kt}$    & minimum number of total remaining flight time for cluster $k$ at period $t$. & hours. \\
            $Rft^{Init}_i$      & remaining flight time for aircraft $i$ at the start of the planning horizon.  & hours.  \\
            $Rct^{Init}_i$      & remaining calendar time for aircraft $i$ at the start of the planning horizon.  & periods. \\
        \end{tabular}

@@

    \begin{align}
        & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
        & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
        & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag
    \end{align}

@@

    \begin{tabular}{p{8mm}p{147mm}}
        $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ at the beginning of period $t$ and finishes it at the end of period $t'$. \\  
        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
        $h_{itt'}$  &  Continuous. Sum of the monthly remaining flight hours for aircraft $i \in I$ for periods between the beginning of period $t$ and the beginning of period $t'$. \\
    \end{tabular}

@@

    \begin{align}
        & \text{Min}\; \sum_{\substack{i \in \mathcal{I}, \\ t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}} m_{it_1t_2} \times H^{max} \times (t_2 - t_1) - h_{it_1t_2}  \label{eq:objective1}
    \end{align}

@@

    \begin{align}
       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}
       % & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
       %  &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
    \end{align}

    Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$. A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission. 
    % Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.

    \begin{align}
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
    \end{align}

@@

    \begin{align}
        & h_{i1t_1} \geq (t_1 - 1) Rft^{Init}_i m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} - U^{min}_{1t_1} 
            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours1}\\
        & h_{it_1t_2} \geq (t_2 - t_1) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} - U^{min}_{t_1t_2}
            &t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours2}\\
        & h_{it_2|\mathcal{T}|} \geq (|\mathcal{T}|- t_2) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} - U^{min}_{t_2|\mathcal{T}|}
            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours3}
    \end{align}

@@

    \begin{align}
       % For now, we only want one or possibly two maintenances...
        & \sum_{\substack{t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}|}} m_{it_1t_2} =  1 
          & i \in \mathcal{I}\label{eq:num_maint} \\
        % & m_{it'} + m_{it} \leq 1
        %   & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\  
        % & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
        %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
        % & m_{it_1t_2} = 0
        %   & t_1 \in \mathcal{T}^{m_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
        & m_{it_1t_2} =  0
          & t_1 \notin \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
        & m_{it_1t_2} =  0
          & t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup |\mathcal{T}|, i \in \mathcal{I} \label{eq:maint_forbid2}
    \end{align}

>>>>

        \begin{tabular}{p{15mm}p{140mm}}
            $i \in \mathcal{I}$     &  aircraft. Between 15 and 60. \\
            $t \in \mathcal{T}$     &  time periods included in the planning horizon. Between 90 and 120 months. \\
            $j \in \mathcal{J}$     &  missions. There are between 1 and 4 active missions at any given time. \\
        \end{tabular}

    \subsubsection{Auxiliary sets}

        \begin{tabular}{p{15mm}p{140mm}}
            $k \in \mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
            $n \in \mathcal{N}$     &  maintenance cycles for aircraft. Up to 3. \\
            $s \in \mathcal{S}$     &  slots for elastic variables. \\
        \end{tabular}

@@

        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
            $N_t$               & number of aircraft known to be in maintenance in period $t$. & aircraft. \\
            $N^{Clust}_{kt}$    & number of aircraft in cluster $k$ known to be in maintenance in period $t$. & aircraft. \\
            $A^{Clust}_{kt}$    & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$. & aircraft. \\
            $H^{Clust}_{kn}$    & maximum number of mean accumulated flight time for cluster $k$ at maintenance cycle $n$. & hours. \\
            $Rft^{Init}_i$      & remaining flight time for aircraft $i$ at the start of the planning horizon.  & hours.  \\
            $Rct^{Init}_i$      & remaining calendar time for aircraft $i$ at the start of the planning horizon.  & periods. \\
        \end{tabular}

    \subsubsection{Objective function weights}

        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
            $PA^{Clust}_s$   & unit penalty for violating serviceability constraint in slot $s$. & \\
            $PH^{Clust}_s$   & unit penalty for violating sustainability constraint in slot $s$. & \\
        \end{tabular}

@@

    \begin{align}
        & U^{\prime}_{t_1t_2} = U^{min} (t_2 - t_1+1) \notag \\
        & U^{\prime\prime}_{t_1t_2} = U^{min} \sum_{t=t_1-1}^{t_2} (t_2 - t) = U^{min} (t_2 - t_1+1) (t_2 - t_1) \div 2  \notag \\
        & H^\prime_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1+1) \notag \\
        & H^{\prime\prime}_{jt_1t_2t'} = (H_j- U^{min}) \sum_{t=t_1-1}^{t_2} (t' - t) = (H_j- U^{min}) (t_2 - t_1 + 1) (2 t' - t_1 - t_2) \div 2 \notag
    \end{align}

@@

    \begin{tabular}{p{8mm}p{147mm}}
        $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ at the beginning of period $t$ and finishes it at the end of period $t'$. \\  
        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
    \end{tabular}

    Auxiliary variables control the status of each aircraft or group of aircraft.

    \begin{tabular}{p{8mm}p{147mm}}
        $h_{in}$  &  Continuous, positive. Sum of the accumulated monthly consumption in flight hours for aircraft $i \in I$ for maintenance cycle $n$. \\
        $e^{AClust}_{kts}$ & Continuous, positive. Elastic variable for serviceability constraint. \\
        $e^{HClust}_{kts}$ & Continuous, positive. Elastic variable for sustainability constraint. \\
    \end{tabular}

@@

    \begin{align}
        & \text{Min}\; 
        \sum_{\substack{i \in \mathcal{I}, \\ 
            t_1 \in \mathcal{T}^{M_{Init}}_i, \\ 
            t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}
            }  
            m_{it_1t_2} \times H^{max} 
        + \sum_{k \in \mathcal{K}, t \in \mathcal{T}, s \in \mathcal{S}} PA^{Clust}_s e^{AClust}_{kts} 
        + \sum_{k \in \mathcal{K}, n \in \mathcal{N}, s \in \mathcal{S}} PH^{Clust}_s e^{HClust}_{kns}
        \label{eq:objective1}
    \end{align}

@@

    \begin{align}
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
    \end{align}

@@

    \begin{align}
        & h_{i1} (t_1 - 1) \geq (t_1 - 1) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} + U^{\prime\prime}_{1t_1} 
            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours1}\\
        & h_{i2} (t_2 - t_1) \geq (t_2 - t_1) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} + U^{\prime\prime}_{t_1t_2}
            &t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours2}\\
        & h_{i3} (|\mathcal{T}|- t_2) \geq (|\mathcal{T}|- t_2) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} + U^{\prime\prime}_{t_2|\mathcal{T}|}
            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours3}
    \end{align}

    \begin{align}
       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt} + \sum_{s \in \mathcal{S}} e^{AClust}_{kts}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}\\
       & \sum_{i \in \mathcal{IK}_k} h_{in} \leq |{IK}_k| H^{Clust}_{kn} + \sum_{s \in \mathcal{S}} e^{HClust}_{kns}
        &k \in \mathcal{K}, n \in \mathcal{N} \label{eq:sustainability-cluster}
    \end{align}

    Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$. A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission. 
    Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.

@@

    \begin{align}
       % For now, we only want one or possibly two maintenances...
        & \sum_{\substack{t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}|}} m_{it_1t_2} =  1 
          & i \in \mathcal{I}\label{eq:num_maint} \\
        & m_{it_1t_2} =  0
          & t_1 \notin \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
        & m_{it_1t_2} =  0
          & t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup |\mathcal{T}|, i \in \mathcal{I} \label{eq:maint_forbid2}
    \end{align}
